---
title: "PS270 Final Project"
author: "Thomas-J-Griffiths"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---
## Introduction

During the 2024 general election in the UK, Reform UK (formally the Brexit Party) ran on a platform of limiting immigration and removing Britain from certain European institutions which would appeal to areas that voted to leave the EU in the Brexit Referendum while discouraging people that voted "Remain" from voting for them as it would go against their preferences stay more aligned with the EU. It would also appeal to blue-collar areas in northern England whom many felt

Is there a relationship between how people voted in the 2016 Brexit Referendum and how much people voted for Reform UK in the 2024 UK general election? For this question, I hypothesis that constituencies and regions that had a higher vote share for leave in the Brexit Referendum had a higher vote share for the Reform UK in the 2024 general election and that regions that had a higher mean leave vote share will have a higher mean Reform UK vote share.

This question is important because the Brexit referendum was a hugely consequential election in the UK and Europe as a whole, beginning the process of the UK leaving the European Union. As such, its important to study how voters the voted leave in the referendum voted in subsequent elections to get a better understanding of the British electorate and how Brexit could still be affecting British politics.  

## Data Organization

The data for this project came from Dr. Henretty, who created a model that estimated how each constituency (post-2023 borders) voted in the Brexit referendum. The 2024 general election data came from the House of Commons Library which collecting all the voting data for each constituency by adding up the voting data from each precinct in all respective constituency. The independent variable for this project are the elections, specifically between the 2016 referendum and the 2024 general election. The dependent variable is how each constituency voted in each party and in particular the party vote share for Reform UK. The study has a difference-in-difference desgin as it is analyzing data collected from multiple points in time, in this case being the referendum in 2016 and the general election in 2024. It isn't a before-and-after degin because we aren't introducing an intervening treatment varaible. This project will only look at constituencies in England and Wales as the party systems in Scotland and Northern Ireland have significant differences. The explanatory variable for this project are the constituencies of England and Wales. The outcome variable is the vote share for leave in the Brexit Referendum and Reform UK vote share in each constituency and region in the 2024 general election. 

The hypothesis would be supported if there is a positive correlation between the vote share of leave in the Brexit Referendum and the vote share of Reform UK and in the 2024 general election and if the regions that had higher leave vote share also have a higher Reform vote share.


## Packages Used 

```{r}
library(tidyverse) 
library(infer)
library(broom)
```

## Assinging Data Sets

Reading Data Sets

```{r}
X2024_Brexit_Results <- read_csv("data/2024_Brexit_Results.csv")

HoC_GE2024_results_by_constituency <- read_csv("data/HoC-GE2024-results-by-constituency.csv")


UK_election_results <- HoC_GE2024_results_by_constituency

Brexit_2024_results <- X2024_Brexit_Results
```
Filtering to English and Welsh Constituencies

```{r}
ENG_WLS_results <- UK_election_results |>
  filter(`Country name` %in% c("England", "Wales"))

ENG_WLS_results <- ENG_WLS_results |>
  rename(Constituen = `Constituency name`)

ENG_WLS_results <- ENG_WLS_results |>
  left_join(Brexit_2024_results)

ENG_WLS_results <- ENG_WLS_results |>
  mutate(
    Con_share = Con / `Valid votes`,
Reform_share = RUK / `Valid votes`,
LD_share = LD / `Valid votes`,
PC_share = PC / `Valid votes`,
Green_share = Green / `Valid votes`,
Lab_share = Lab / `Valid votes`
)
        
```

## Data

2024 Reform Vote Share and Vote Leave Plot for England and Wales

```{r}
ENG_WLS_results |>
  filter(Reform_share > 0) |>
ggplot(data = ENG_WLS_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (England and Wales)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

2024 Reform Vote Share and Vote Leave Histogram for England and Wales

```{r}
ENG_WLS_results |>
  filter(Reform_share > 0) |>
  ggplot(mapping = aes(x = LeavePct)) +
  geom_histogram()
```

2024 Reform Vote Share and Vote Leave Plots by Region

Northeast

```{r}
NE_results <- ENG_WLS_results |>
  filter(`Region name` == "North East")

NE_results |>
  filter(Reform_share > 0) |>
ggplot(data = NE_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (North East England)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

Northwest

```{r}
NW_results <- ENG_WLS_results |>
  filter(`Region name` == "North West")

NE_results |>
  filter(Reform_share > 0) |>
ggplot(data = NW_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (North West England)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

Yorkshire and The Humber

```{r}
Yorkshire_results <- ENG_WLS_results |>
  filter(`Region name` == "Yorkshire and The Humber")

Yorkshire_results |>
  filter(Reform_share > 0) |>
ggplot(data = Yorkshire_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (Yorkshire and Humber)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

East Midlands

```{r}
Eastmid_results <- ENG_WLS_results |>
  filter(`Region name` == "East Midlands")

Eastmid_results |>
  filter(Reform_share > 0) |>
ggplot(data = Eastmid_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (East Midlands)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

West midlands

```{r}
Westmid_results <- ENG_WLS_results |>
  filter(`Region name` == "West Midlands")

Westmid_results |>
  filter(Reform_share > 0) |>
ggplot(data = Westmid_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (West Midlands)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

East of England

```{r}
Easteng_results <- ENG_WLS_results |>
  filter(`Region name` == "East of England")

Easteng |>
  filter(Reform_share > 0) |>
ggplot(data = Easteng_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (East of England)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

South West

```{r}
SW_results <- ENG_WLS_results |>
  filter(`Region name` == "South West")

SW_results |>
  filter(Reform_share > 0) |>
ggplot(data = SW_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (South West)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

South East

```{r}
SE_results <- ENG_WLS_results |>
  filter(`Region name` == "South East")

SE_results |>
  filter(Reform_share > 0) |>
ggplot(data = SE_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (South East)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

London

```{r}
London_results <- ENG_WLS_results |>
  filter(`Region name` == "London")

London_results |>
  filter(Reform_share > 0) |>
ggplot(data = London_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (London)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

Wales

```{r}
WLS_results <- ENG_WLS_results |>
  filter(`Region name` == "Wales")

WLS_results |>
  filter(Reform_share > 0) |>
ggplot(data = WLS_results, mapping = aes(x = LeavePct, y = Reform_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Reform UK 2024 Vote Share",
  title = "Brexit Referendum and Reform 2024 Vote Share (Wales)"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```



Multiple Regression Model



```{r}
Reform_multreg <- lm(LeavePct ~ Reform_share + Lab_share + LD_share + Green_share + Con_share, data = Yorkshire_results)

 tidy(Reform_multreg) |>
     knitr::kable(digits = 3)
```

```{r}
Region_multreg <- lm(LeavePct ~ Reform_share + Con_share + Green_share + LD_share + Lab_share, data = ENG_WLS_results)
tidy(Region_multreg) |>
     knitr::kable(digits = 3)
```


Mean Vote Share

```{r}
means_table <- ENG_WLS_results |>
 group_by(`Region name`) |>
  summarize(mean_LeavePct = mean(LeavePct, na.rm = TRUE),
            mean_Reform_share = mean(Reform_share)) |>
  rename("Region" = `Region name`,
         "Mean Leave Vote Share" = mean_LeavePct,
         "Mean Reform Vote Share" = mean_Reform_share) |>
     knitr::kable(digits = 3) 
 
```


## Data Interpretation

We can interpret from the regression model, that we can very confident that there is a relationship between vote leave percentage and the vote share for Reform UK. The regression model shows a p-value of approximately 0 for Reform vote share and when holding all other major parties vote share constant. Having a coefficient of around 0.9 indicating that indicating that for every percent increase in vote leave, the Reform vote share will increase by an average of around 0.9%.

## Conlusion

In conclusion, the data supports the hypothesis that there is a positive correlation between Reform UK and the Conservative vote share with leave vote share. It also supports the hypothesis that there is a negative correlation between areas that vote leave and Liberal Democrats and Greens. The hypothesis was most supported by the Greens and Reform data as they had the strongest relationships with vote leave. The hypothesis that Labour vote share wouldn't have a meaningful relationship with vote lave was not supported by the data as they had a statically significant negative correlation with vote leave. If I had more resources and time, I would add in data from the 2019 general election to see if there is there any patterns in party vote share and vote leave. Also, I would try to get data on how each country/region voted to see if there are any regional differences for the relationship between how people voted in the Brexit Referendum and the general election.


## Apendix

The Conservatives ran on similar issues to Reform UK on reducing immigration with some members arguing for the withdraw from European institutions, although to a lesser extent and more moderate. Many of their members also campaigned for leaving the EU during the referendum, which, like Reform UK, would appeal to areas that largely voted to leave the EU.

The Liberal Democrats and Greens, on the other hand, campaigned for Remain during the referendum and during the 2024 election supported the UK rejoining EU institutions and eventually rejoining the EU. These party positions would have the opposite effect of Reform UK and the Conservatives in constituencies that largely voted for leave, with those constituencies not having much support for the Liberal Democrats. Instead, the Liberal Democrats and Greens pro-EU policies would appeal to areas that had a lower share that voted leave.

From the multiple regression model, it appears that the correlation between Reform UK vote share and vote leave percent is stronger than the relationship between the Conservative vote share and vote leave percent. This is apparent from the Reform share coefficient being around 1.16 which indicates that the average increase in Reform vote share is around 1.16% for each percent increase in vote leave share holding Conservative vote share fixed. In comparison, the Conservative vote share coefficient is only around 0.22, indicating that the average increase in Conservative vote share is around 0.22% for each percent increase in the share for vote leave holding Reform UK vote share fixed. Reform UK having a stronger positive relationship can be seen in the plots as clear positive curve while the Conservative plot has a clear positive slope until around the 50% vote leave mark where it then remains mostly flat.

For the Liberal Democrats and Green multiple regression, it appears that there is a negative correlation between their vote share and leave vote share. The correlation appears to be stronger for the Greens, having a coefficient around -1.05, indicating that the average decrease in Green vote share is around 1.05% for every percent increase in leave vote share holding Liberal Democrat vote share. The Liberal Democrats have a weaker, but still significant coefficient of -0.304, indicating that the average decrease in Liberal Democratic vote share is around 0.304% for every percent increase in leave vote share holding Green vote share. This relationship is apparent in the plots as Green plot appears to have a more consistent negative slope compared to the Liberal Democrat one, which, while having an overall negative slope, has more of a bell-shaped curve.


2024 Conservative Vote Share and Vote Leave Plot

```{r}
ENG_WLS_results |>
  filter(Con_share > 0) |>
ggplot(data = ENG_WLS_results, mapping = aes(x = LeavePct, y = Con_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Conservative Party 2024 Vote Share",
  title = "Brexit Referendum and 2024 Conservative Party Vote Share England and Wales"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

2024 Labour Vote Share and Vote Leave Plot

```{r}
ENG_WLS_results |>
   filter(Lab_share > 0) |>
ggplot(data = ENG_WLS_results, mapping = aes(x = LeavePct, y = Lab_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Labour 2024 Vote Share",
  title = "Brexit Referendum and Labour 2024 Vote Share England and Wales"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

2024 Liberal Democrats Vote Share and Vote Leave Plot

```{r}
ENG_WLS_results |>
   filter(LD_share > 0) |>
ggplot(data = ENG_WLS_results, mapping = aes(x = LeavePct, y = LD_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Liberal Democrats 2024 Vote Share",
  title = "Brexit Referendum and Liberal Democrat 2024 Vote Share England and Wales"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

2024 Green Vote Share and Vote Leave Plot

```{r}
ENG_WLS_results |>
   filter(Green_share > 0) |>
ggplot(data = ENG_WLS_results, mapping = aes(x = LeavePct, y = Green_share)) +
  geom_point(color = "turquoise4", alpha = .5) +
  geom_smooth(color = "indianred", se = FALSE) +
  labs(
    x = "Vote Leave Percentage",
    y = "Green 2024 Vote Share",
  title = "Brexit Referendum and Green 2024 Vote Share England and Wales"
  ) +
   annotate(
    "text", 
    x = Inf, y = -Inf, 
    label = "Sources: Dr Henretty, House of Commons Library", 
    hjust = 1, vjust = -.25, 
    size = 3, color = "gray20", fontface = "italic"
  ) +
  theme_minimal()
```

Multiple Regression Model for Other Parties

```{r}
Reform_Con_multreg <- lm(LeavePct ~ Reform_share + Con_share, data = ENG_WLS_results)
  tidy(Reform_Con_multreg) |>
     knitr::kable(digits = 3)
```

Sources

House of Commons Library, General Election 2024 Results
https://commonslibrary.parliament.uk/research-briefings/cbp-10009/

Dr. Henretty, 2016 Brexit referendum estimates on 2024 boundaries
https://docs.google.com/spreadsheets/d/1mtph-ml7CYVoeEUIc1g_IbOvbiZMa_ezRGQlHQoCpF4/edit?pli=1&gid=341382343#gid=341382343


```{r}
LD_Green_multreg <- lm(LeavePct ~ LD_share + Green_share, data = ENG_WLS_results)

 tidy(LD_Green_multreg) |>
     knitr::kable(digits = 3)
```